Forest canopies are better generators of biodiversity in the tropics

### 2025-06-30 manuscript.Rmd
canopy height richness models and plots

@@

>>>>

```{r}
# analysis scripts

"scripts/03_analyses/01_canopyheight_richness_relationships.R" # canopy height richness models and plots
```

@@

We classified biomes into larger categories based on the 2017 biome categories. For biomes that occurred across latitudes (e.g., Montane Grasslands and Shrublands), we classified them as tropical if they occurred within the latitudes covered by biomes classified as tropical and temperate if they occurred within the latitudes classified as temperate. Additionally, Mediterranean forests, woodlands, and scrub were partitioned into temperate forests if the ecoregion contained "forest" and temperate woodland, savanna, and shrubland if the ecoregion name did not contain the word "forest".

glmmTMB was used to produce models of canopy height by richness. sdmTMB was used to produce models of verticality. DHARMa was used to check model fit and assumptions.

```{r}
# Fig. 1

# Legend: a) canopy height, b) vertebrate richness (all vertebrates combined)
# c) Deviance residuals for a model of total vertebrate richness by canopy height
# d) Relationships between species richness and canopy height for each taxon in each
# biome category. Dots indicate observed values and lines indicate negative binomial
# model fits of richness by canopy height for each combination of taxon and biome.

"figures/ms_figures/fig1_canopy_richness.png"

```

# Data availability statement
canopy height:
